---
title: "Public opinion poll analysis - Sri Lankan presidential election 2019 "
output: html_notebook
---

This is a breif analysis of the sample opinion poll for the upcoming presidential election in Sri Lanka. The poll was conducted using a questionnaire provided to random individuals using both door to door and online strategies. The sample includes 206 entries covering multiple ethinicities and electoral districts around sri lanka. 

### contributors

1. Randika Senanayake  -
2. Wishwa  Perera      - 199355V

#### Loading the data

```{r}
library(readxl)
responses <- read_excel("questionnaire/responses/responses.xlsx")
#View(responses)
nrow(responses)
```

### Cleaning the data

let us find which columns include incomplete data.
```{r}
colnames(responses)[colSums(is.na(responses))>0]
```
let us remove the rows with incomplete data
```{r}
complete_responses <- na.omit(responses)
nrow(complete_responses)
```

### Visualizations

#### Intent to vote (Yes/No/Dont know/ Other)

```{r}
q1d = as.data.frame(complete_responses[,2],drop=FALSE)
q1dt <- as.data.frame(table(q1d))
#View(q1dt)
levels(q1dt$q1d)
levels(q1dt$q1d) = c("1","2","4","3")
q1dt <- rbind(q1dt,c(3,0))
#View(q1dt)
q1l <- c("Yes","No","prefer not to answer","No idea")
piepercent<- round(100*q1dt[,2]/sum(q1dt[,2]), 1)
pie(q1dt[,2], labels = piepercent, main = "Intent to vote",col = rainbow(length(q1dt[,2])))
legend("topright", q1l, cex = 0.8,fill = rainbow(length(q1dt[,2])))
```
 
#### Party Preference 

```{r}
q2d = as.data.frame(complete_responses[,3],drop=FALSE)
q2dt <- as.data.frame(table(q2d))
#View(q2dt)
q2l <- c("United National Party","Sri Lanka Podujana Peremuna","National People's Movement","National Peoples Party","No Idea","Prefer not to answer","Other")
piepercent2 <- round(100*q2dt[,2]/sum(q2dt[,2]), 1)
pie(q2dt[,2], labels = piepercent2, main = "Party Preference ",col = rainbow(length(q2dt[,2])))
legend("topright", q2l, cex = 0.5,fill = rainbow(length(q2dt[,2])))
```

#### Candidate Preference

```{r}
q3d = as.data.frame(complete_responses[,4],drop=FALSE)
q3dt <- as.data.frame(table(q3d))
#View(q3dt)
q3l <- c("Sajith Premadasa","Gotabhaya Rajapaksha","Anura Kumara Dissanayake","Mahesh Senanayake","Rohan Pallewatte","No idea","Prefer not to answer","Other")
piepercent3 <- round(100*q3dt[,2]/sum(q3dt[,2]), 1)
pie(q3dt[,2], labels = piepercent3, main = "Candidate Preference",col = rainbow(length(q3dt[,2])))
legend("topright", q3l, cex = 0.5,fill = rainbow(length(q3dt[,2])))
```
#### Most improtant issue to be addressed

```{r}
q4d = as.data.frame(complete_responses[,5],drop=FALSE)
q4dt <- as.data.frame(table(q4d))
#View(q4dt)
q4dt_mod <- q4dt[1:7,]
#View(q4dt_mod)
q4dt_mod <- rbind(q4dt_mod,c(8,8))
q4l <- c("National Security","Living Cost","Economic Growth","Civil and religious conflicts","Foreign interference","Unemployment","Other")
piepercent4 <- round(100*q4dt_mod[,2]/sum(q4dt_mod[,2]), 1)
pie(q4dt_mod[,2], labels = piepercent4, main = "Most improtant issue to be addressed",col = rainbow(length(q4dt_mod[,2])))
legend("topright", q4l, cex = 0.5,fill = rainbow(length(q4dt_mod[,2])))
```


#### Last election vote

```{r}
q5d = as.data.frame(complete_responses[,6],drop=FALSE)
q5dt <- as.data.frame(table(q5d))
#View(q5dt)
q5dt_mod <- q5dt[1:4,]
#View(q5dt_mod)
q5dt_mod <- rbind(q5dt_mod,c(5,3))
q5l <- c("MaithreePala Sirisena","Mahinda Rajapaksha","Did not vote","Prefer not to answer","Other")
piepercent5 <- round(100*q5dt_mod[,2]/sum(q5dt_mod[,2]), 1)
pie(q5dt_mod[,2], labels = piepercent5, main = "Last election vote",col = rainbow(length(q5dt_mod[,2])))
legend("topright", q5l, cex = 0.5,fill = rainbow(length(q5dt_mod[,2])))
```

#### Voters by age

```{r}
q6d = as.data.frame(complete_responses[,7],drop=FALSE)
q6dt <- as.data.frame(table(q6d))
#View(q6dt)
levels(q6dt$q6d) = c("1","2","3","4","5","6","8","9","7","10","11")
q6dt <- rbind(q6dt,c(7,0))
q6dt <- rbind(q6dt,c(10,0))
q6dt <- rbind(q6dt,c(11,0))
#View(q6dt)
q6l <- c("18-23","24-29","30-35","36-41","42-47","48-53","60-65","66-71","54-59","72 above","Prefer not to answer")
barplot( round(100*q6dt[,2]/sum(q6dt[,2]), 1),main="Voters by age",xlab="Age",ylab = "Percentage",names.arg  = q6l, axes=TRUE,las=2,cex/cex.axis = 0.5)
```

#### Voters by gender

```{r}
q7d = as.data.frame(complete_responses[,8],drop=FALSE)
q7dt <- as.data.frame(table(q7d))
#View(q7dt)
levels(q7dt$q7d) = c("1","2","4","3")
q7dt <- rbind(q7dt,c(3,0))
q7l <- c("Female","Male","Other","Prefer not to answer")
piepercent7 <- round(100*q7dt[,2]/sum(q7dt[,2]), 1)
pie(q7dt[,2], labels = piepercent7, main = "Voters by gender",col = rainbow(length(q7dt[,2])))
legend("topright", q7l, cex = 0.5,fill = rainbow(length(q7dt[,2])))
```

#### Voters by ethnicity

```{r}
q8d = as.data.frame(complete_responses[,9],drop=FALSE)
q8dt <- as.data.frame(table(q8d))
#View(q8dt)
q8l <- c("Sinhala","Tamil","Muslim","Bhurger","Prefer Not to answer","Other")
piepercent8 <- round(100*q8dt[,2]/sum(q8dt[,2]), 1)
pie(q8dt[,2], labels = piepercent8, main = "Voters by ethnicity",col = rainbow(length(q8dt[,2])))
legend("topright", q8l, cex = 0.5,fill = rainbow(length(q8dt[,2])))
```


#### Voters by religion

```{r}
q9d = as.data.frame(complete_responses[,10],drop=FALSE)
q9dt <- as.data.frame(table(q9d))
#View(q9dt)
q9l <- c("Buddhist","Hindu","Christian","Islam","Prefer Not to answer","Other")
piepercent9 <- round(100*q9dt[,2]/sum(q9dt[,2]), 1)
pie(q9dt[,2], labels = piepercent9, main = "Voters by ethnicity",col = rainbow(length(q9dt[,2])))
legend("topright", q9l, cex = 0.5,fill = rainbow(length(q9dt[,2])))
```

#### Voters by Education

```{r}
q10d = as.data.frame(complete_responses[,11],drop=FALSE)
q10dt <- as.data.frame(table(q10d))
#View(q10dt)
levels(q10dt$q10d) = c("4","5","1","2","3")
q10dt <- rbind(q10dt,c(1,0))
q10dt <- rbind(q10dt,c(2,0))
q10dt <- rbind(q10dt,c(3,0))
q10l <- c("Degree/Post Graduate","Prefer not to answer","Illiterate","Elementary Education","O/L","A/L")
piepercent10 <- round(100*q10dt[,2]/sum(q10dt[,2]), 1)
pie(q10dt[,2], labels = piepercent10, main = "Voters by Education",col = rainbow(length(q10dt[,2])))
legend("topright", q10l, cex = 0.5,fill = rainbow(length(q10dt[,2])))
```

#### Voters by occupation

```{r}
q11d = as.data.frame(complete_responses[,12],drop=FALSE)
q11dt <- as.data.frame(table(q11d))
#View(q11dt)
levels(q11dt$q11d) = c("1","2","3","4","5","6","8","9","7","10","11")
q11l <- c("Student","Private Sector","Government Sector","Self employed/ Business owner","Unemployed","Prefer not to answer")
barplot( round(100*q11dt[,2]/sum(q11dt[,2]), 1),main="Voters by occupation",xlab="Occupation",ylab = "Percentage",names.arg  = q11l, axes=TRUE,las=2,cex.names = 0.5)
```

### Voters by Income
```{r}
q12d = as.data.frame(complete_responses[,13],drop=FALSE)
q12dt <- as.data.frame(table(q12d))
q12dt_mod <- q12dt[1:6,]
q12dt_mod <- rbind(q12dt_mod,c(7,3))
q12l <- c("0-20k","20k-50k","50k-100k","100k-200k","200k-300k","Prefer not to answer","other")
barplot( round(100*q12dt_mod[,2]/sum(q12dt_mod[,2]), 1),main="Voters by Income",xlab="Income in Rupees",ylab = "Percentage",names.arg  = q12l, axes=TRUE,las=2,xpd=FALSE, cex.names = 0.5)
```