---
title: "Public opinion poll analysis - Sri Lankan presidential election 2019 "
output: html_notebook
---

This is a breif analysis of the sample opinion poll for the upcoming presidential election in Sri Lanka. The poll was conducted using a questionnaire provided to random individuals using both door to door and online strategies. The sample includes 206 entries covering multiple ethinicities and electoral districts around sri lanka. 

### contributors

1. Randika Senanayake  -
2. Wishwa  Perera      - 199355V

#### Loading the data

```{r}
library(readxl)
responses <- read_excel("questionnaire/responses/responses.xlsx")
View(responses)
nrow(responses)
```

### Cleaning the data

let us find which columns include incomplete data.
```{r}
colnames(responses)[colSums(is.na(responses))>0]
```
let us remove the rows with incomplete data
```{r}
complete_responses <- na.omit(responses)
nrow(complete_responses)
```

### Visualizations

#### Intent to vote (Yes/No/Dont know/ Other)

```{r}
q1d = as.data.frame(complete_responses[,2],drop=FALSE)
q1dt <- as.data.frame(table(q1d))
View(q1dt)
levels(q1dt$q1d)
levels(q1dt$q1d) = c("1","2","4","3")
q1dt <- rbind(q1dt,c(3,0))
View(q1dt)
q1l <- c("Yes","No","No idea","prefer not to answer")
piepercent<- round(100*q1dt[,2]/sum(q1dt[,2]), 1)
pie(q1dt[,2], labels = piepercent, main = "Intent to vote",col = rainbow(length(q1dt[,2])))
legend("topright", q1l, cex = 0.8,fill = rainbow(length(q1dt[,2])))
```
 
#### Winning party 

```{r}
q2d = as.data.frame(complete_responses[,3],drop=FALSE)
q2dt <- as.data.frame(table(q2d))
View(q2dt)
q2l <- c("United National Party","Sri Lanka Podujana Peremuna","National People's Movement","National Peoples Party","No Idea","Prefer not to answer","Other")
piepercent<- round(100*q2dt[,2]/sum(q2dt[,2]), 1)
pie(q2dt[,2], labels = piepercent, main = "Winning Party",col = rainbow(length(q2dt[,2])))
legend("topright", q2l, cex = 0.5,fill = rainbow(length(q2dt[,2])))
```
